Sex and Power: Sexual dimorphism in trait variability and its eco-evolutionary and statistical implications

### 2020-05-18 scripts/MouseSexDiffApril20_2020.Rmd
More major updates to Supplement, reorganising, linking with fig 1.1 and cutting down redundant code

Our supplementary material follows the general structure described in our methods (Figure  \@ref(fig:FigG1)). Generally speaking, the first few sections describe the processes by which the data were cleaned and re-organised for analysis, and how the first order and second order meta-analsyes were conducted. Supplemental tables and figures that are referenecd within the main manuscript can be found at the end of the document or conviently referenecd quickly using the side table of contents panel.

@@

```{r}
data_subset_parameterid_individual_by_age <- function(mydata, parameter, age_min=0, age_center=100) {
  tmp <- mydata %>%
    filter(
      age_in_days >= age_min,
      id == parameter
    ) %>%
    # take results for single individual closest to age_center
    mutate(age_diff = abs(age_center - age_in_days)) %>%
    group_by(biological_sample_id) %>%
    filter(age_diff == min(age_diff)) %>%
    select(-age_diff)# %>%

  # still some individuals with multiple records (because same individual appear under different procedures, so filter to one record)
  j <- match(unique(tmp$biological_sample_id), tmp$biological_sample_id)
  tmp[j, ] 
  }
```

### Population statistics
Create a function called: "calculate_population_stats"
This function groups animals from the same strain and institution together. This is done for each trait separately, and only for traits that have been measured in both sexes. Any group containing fewer than 5 individuals is excluded.

multi_spread <- function(df, key, value) {
    # quote key
    keyq <- rlang::enquo(key)
    # break value vector into quotes
    valueq <- rlang::enquo(value)
    s <- rlang::quos(!!valueq)
    df %>% gather(variable, value, !!!s) %>%
        unite(temp, !!keyq, variable) %>%
        spread(temp, value)
}

# Will calculate population stats and exclude data that is irrelevant and not useful, it will also just convert the data straight away to wide format as opposed to long, making it ready for effect size calculations

### Extraction of effect sizes and sample variances
Function: "create_meta_analysis_effect_sizes"

#This function should mostly replace the entire function below, but I am still not quite sure what is happening in places. Adding all effect sizes together (maintaining the naming), err seems to be added to each trait? That needs to be added.It will make use of escalc which has the correct equations for effect size calculations.
create_meta_analysis_effect_sizes <- function(mydata, measure = c("CVR", "VR", "ROM")){
  for(i in 1:length(measure)){
    mydata <- metafor::escalc(m1i  = male_x_bar, 
                              m2i  = female_x_bar, 
                              sd1i = male_x_sd, 
                              sd2i = female_x_sd, 
                              n1i  = male_n_ind, 
                              n2i  = female_n_ind, 
                              data = mydata, 
                           measure = measure[i], 
                         var.names = c(paste0("effect_size_", measure[i]), 
                                      paste0("sample_variance_", measure[i])),
                            append = TRUE)
    }
    return(mydata)
}
```

## Load & clean data

### Data loading and cleaning of the csv file

We have already done this step and provide a cleaned up file which is less computationally intensive to deal with. The file has been saved in a folder called `export`. However, the `.csv` is provided in case this is preferred and can be imported and processed using following the steps below: [This code is currently excluded from showing in this `.html` file, but can be viewed and enabled / run in the down-loadable .rmd file. However, it requires the .gz file containing the raw data in the "export" folder. Unfortunately, that file can not be provided via Github, as it is too large (274MB). However, it is freely accessible and uploaded to zenodo.org (https://zenodo.org/deposit/3759701)

```{r clean, eval=FALSE, include=FALSE}
# loads the raw data, setting some default types for various columns

@@

# Load raw data - save cleaned dataset as RDS for reuse
data_raw <- load_raw(here("data","dr7.0_all_control_data.csv.gz"))
dir.create("export", F, F)

@@

# Meta-analyses
## Population as analysis unit 
(Step C, Figure 3 in main document)

@@

```{r loop, message=FALSE, warning=FALSE, include=FALSE}

@@

## Meta-analysis: condensing non-independent traits 
(Step F in Figure 3 in main article)

@@

## Second-order meta-analysis for functional groups
(Section H in Figure 3 in main article)

### Performing meta-analyses 
Three for each of the 9 grouping terms: lnCVR, lnVR, lnRR
#### Preparing the data

@@

# Visualising meta-analysis results
## Preparation for Figure 4 
#### Count data, based on First-order meta-analysis results
This includes all separate eligible traits.
Re-ordering of grouping terms. 
Preparation of subplot 4A

@@

#### Overall results of second order meta-analysis (Figure 4, Panel B)
#### Re-structure data for plotting 
Data are re-structured, and grouping terms are being re-ordered

@@

Panel A shows the numbers of traits across functional groups that are either male-biased (blue-green) or female-biased (orange-red), as calculated in Step D (figure 3). Panel B shows effect sizes and 95% CI from separate meta-analysis for each functional group (step H in Figure 3). Both panels represent results evaluated across all traits (Phase 3, Figure 3). Traits that are male biased are shown in blue, whereas female bias data is represented in orange.

@@

## Preparation for Figure 5
#### Preparing data for traits with CI not overlapping 0
To further investigate sex bias in this dataset, and in particular if the extent of sex bias differs between traits, we investigate the magnitude of male- and female bias in significantly different traits on (both for means and variability)

@@

#### Figure 5 B - traits with CI not overlapping 0 
# Prepare data 
# create column with 1= different from zero, 0= zero included in CI

@@

# as above: Prepare data for traits with CI not overlapping 0
# create column with 1= different from zero, 0= zero included in CI

@@

plot3.female.meta.VR <- metacombo_female.plot3.VR %>%
  mutate(model_lnVR = map(data, ~ metafor::rma.uni(
    yi = .x$lnVR, sei = (.x$lnVR_upper - .x$lnVR_lower) / (2 * 1.96),
    control = list(optimizer = "optim", optmethod = "Nelder-Mead", maxit = 1000), verbose = F
  )))

plot3.female.meta.RR <- metacombo_female.plot3.RR %>%
  mutate(model_lnRR = map(data, ~ metafor::rma.uni(
    yi = .x$lnRR, sei = (.x$lnRR_upper - .x$lnRR_lower) / (2 * 1.96),
    control = list(optimizer = "optim", optmethod = "Nelder-Mead", maxit = 1000), verbose = F
  )))

@@

plot3.female.meta.CVR.all <- metacombo_female.plot3.CVR.all %>%
  mutate(model_lnCVR = map(data, ~ metafor::rma.uni(
    yi = .x$lnCVR, sei = (.x$lnCVR_upper - .x$lnCVR_lower) / (2 * 1.96),
    control = list(optimizer = "optim", optmethod = "Nelder-Mead", maxit = 1000), verbose = F
  )))

@@

plot3.female.meta.VR.all <- metacombo_female.plot3.VR.all %>%
  mutate(model_lnVR = map(data, ~ metafor::rma.uni(
    yi = .x$lnVR, sei = (.x$lnVR_upper - .x$lnVR_lower) / (2 * 1.96),
    control = list(optimizer = "optim", optmethod = "Nelder-Mead", maxit = 1000), verbose = F
  )))

@@

plot3.female.meta.RR.all <- metacombo_female.plot3.RR.all %>%
  mutate(model_lnRR = map(data, ~ metafor::rma.uni(
    yi = .x$lnRR, sei = (.x$lnRR_upper - .x$lnRR_lower) / (2 * 1.96),
    control = list(optimizer = "optim", optmethod = "Nelder-Mead", maxit = 1000), verbose = F
  )))

@@

plot3.female.meta.CVR.b <- as.data.frame(plot3.female.meta.CVR %>% group_by(GroupingTerm) %>%
  mutate(
    lnCVR = map_dbl(model_lnCVR, pluck(2)), lnCVR_lower = map_dbl(model_lnCVR, pluck(6)),
    lnCVR_upper = map_dbl(model_lnCVR, pluck(7)), lnCVR_se = map_dbl(model_lnCVR, pluck(3)),
    lnCVR_I2 = map_dbl(model_lnCVR, pluck("I2"))
  ))[, c(1, 4:8)]

@@

plot3.female.meta.VR.b <- as.data.frame(plot3.female.meta.VR %>% group_by(GroupingTerm) %>%
  mutate(
    lnVR = map_dbl(model_lnVR, pluck(2)), lnVR_lower = map_dbl(model_lnVR, pluck(6)),
    lnVR_upper = map_dbl(model_lnVR, pluck(7)), lnVR_se = map_dbl(model_lnVR, pluck(3)),
    lnVR_I2 = map_dbl(model_lnVR, pluck("I2"))
  ))[, c(1, 4:8)]

@@

plot3.female.meta.RR.b <- as.data.frame(plot3.female.meta.RR %>% group_by(GroupingTerm) %>%
  mutate(
    lnRR = map_dbl(model_lnRR, pluck(2)), lnRR_lower = map_dbl(model_lnRR, pluck(6)),
    lnRR_upper = map_dbl(model_lnRR, pluck(7)), lnRR_se = map_dbl(model_lnRR, pluck(3)),
    lnRR_I2 = map_dbl(model_lnRR, pluck("I2"))
  ))[, c(1, 4:8)]

@@

lnCVR.ci <- overall3.female.sig %>%
  filter(parameter == "lnCVR") %>%
  mutate(ci.low = lnCVR_lower, ci.high = lnCVR_upper)
# lnVR.ci <- overall3.female.sig  %>% filter(parameter == "lnVR") %>% mutate(ci.low = lnVR_lower, ci.high = lnVR_upper)
lnRR.ci <- overall3.female.sig %>%
  filter(parameter == "lnRR") %>%
  mutate(ci.low = lnRR_lower, ci.high = lnRR_upper)

@@

## Figure 5
Sex biases in trait values across trait groups (Step J in Figure 3). The panels in A show differences in the number of affected traits, in variance (lnCVR) and means (lnRR), where there is a significant difference between the sexes (i.e CI not overlapping zero). Panel B shows results for the sex bias in those traits that differ between the sexes (second-order meta-analysis). Triangles represent sex bias in means (response ratio) and black circles differences in the coefficient of variation ratio (mean-adjusted variability). The orange-red bars represent trait groups with a female bias, blue-green bars male-biased traits.

@@

  mutate(
    model_heteroCVR = map(data, ~ metafor::rma.uni(
      yi = .x$var.CVR, sei = sqrt(1 / 2 * (.x$N.CVR - 1)),
      control = list(optimizer = "optim", optmethod = "Nelder-Mead", maxit = 10000, stepadj = 0.5), verbose = F
    )),
    model_heteroVR = map(data, ~ metafor::rma.uni(
      yi = .x$var.VR, sei = sqrt(1 / 2 * (.x$N.VR - 1)),
      control = list(optimizer = "optim", optmethod = "Nelder-Mead", maxit = 10000, stepadj = 0.5), verbose = F
    )),
    model_heteroRR = map(data, ~ metafor::rma.uni(
      yi = .x$var.RR, sei = sqrt(1 / 2 * (.x$N.RR - 1)),
      control = list(optimizer = "optim", optmethod = "Nelder-Mead", maxit = 10000, stepadj = 0.5), verbose = F
    ))
  )

@@

  mutate(
    model_heteroCVR = map(data, ~ metafor::rma.uni(
      yi = .x$var.CVR, sei = sqrt(1 / 2 * (.x$N.CVR - 1)),
      control = list(optimizer = "optim", optmethod = "Nelder-Mead", maxit = 10000, stepadj = 0.5), verbose = F
    )),
    model_heteroVR = map(data, ~ metafor::rma.uni(
      yi = .x$var.VR, sei = sqrt(1 / 2 * (.x$N.VR - 1)),
      control = list(optimizer = "optim", optmethod = "Nelder-Mead", maxit = 10000, stepadj = 0.5), verbose = F
    )),
    model_heteroRR = map(data, ~ metafor::rma.uni(
      yi = .x$var.RR, sei = sqrt(1 / 2 * (.x$N.RR - 1)),
      control = list(optimizer = "optim", optmethod = "Nelder-Mead", maxit = 10000, stepadj = 0.5), verbose = F
    ))
  )

@@

Behaviour. <- heterog1 %>%
  filter(., GroupingTerm == "Behaviour") %>%
  select(., -data) %>%
  mutate(
    heteroCVR = .[[2]][[1]]$b, heteroCVR_lower = .[[2]][[1]]$ci.lb, heteroCVR_upper = .[[2]][[1]]$ci.ub, heteroCVR_se = .[[2]][[1]]$se,
    heteroVR = .[[3]][[1]]$b, heteroVR_lower = .[[3]][[1]]$ci.lb, heteroVR_upper = .[[3]][[1]]$ci.ub, heteroVR_se = .[[3]][[1]]$se,
    heteroRR = .[[4]][[1]]$b, heteroRR_lower = .[[4]][[1]]$ci.lb, heteroRR_upper = .[[4]][[1]]$ci.ub, heteroRR_se = .[[4]][[1]]$se
  ) %>%
  select(., GroupingTerm, heteroCVR:heteroRR_se)

Immunology. <- heterog1 %>%
  filter(., GroupingTerm == "Immunology") %>%
  select(., -data) %>%
  mutate(
    heteroCVR = .[[2]][[1]]$b, heteroCVR_lower = .[[2]][[1]]$ci.lb, heteroCVR_upper = .[[2]][[1]]$ci.ub, heteroCVR_se = .[[2]][[1]]$se,
    heteroVR = .[[3]][[1]]$b, heteroVR_lower = .[[3]][[1]]$ci.lb, heteroVR_upper = .[[3]][[1]]$ci.ub, heteroVR_se = .[[3]][[1]]$se,
    heteroRR = .[[4]][[1]]$b, heteroRR_lower = .[[4]][[1]]$ci.lb, heteroRR_upper = .[[4]][[1]]$ci.ub, heteroRR_se = .[[4]][[1]]$se
  ) %>%
  select(., GroupingTerm, heteroCVR:heteroRR_se)

Hematology. <- heterog1 %>%
  filter(., GroupingTerm == "Hematology") %>%
  select(., -data) %>%
  mutate(
    heteroCVR = .[[2]][[1]]$b, heteroCVR_lower = .[[2]][[1]]$ci.lb, heteroCVR_upper = .[[2]][[1]]$ci.ub, heteroCVR_se = .[[2]][[1]]$se,
    heteroVR = .[[3]][[1]]$b, heteroVR_lower = .[[3]][[1]]$ci.lb, heteroVR_upper = .[[3]][[1]]$ci.ub, heteroVR_se = .[[3]][[1]]$se,
    heteroRR = .[[4]][[1]]$b, heteroRR_lower = .[[4]][[1]]$ci.lb, heteroRR_upper = .[[4]][[1]]$ci.ub, heteroRR_se = .[[4]][[1]]$se
  ) %>%
  select(., GroupingTerm, heteroCVR:heteroRR_se)

Hearing. <- heterog1 %>%
  filter(., GroupingTerm == "Hearing") %>%
  select(., -data) %>%
  mutate(
    heteroCVR = .[[2]][[1]]$b, heteroCVR_lower = .[[2]][[1]]$ci.lb, heteroCVR_upper = .[[2]][[1]]$ci.ub, heteroCVR_se = .[[2]][[1]]$se,
    heteroVR = .[[3]][[1]]$b, heteroVR_lower = .[[3]][[1]]$ci.lb, heteroVR_upper = .[[3]][[1]]$ci.ub, heteroVR_se = .[[3]][[1]]$se,
    heteroRR = .[[4]][[1]]$b, heteroRR_lower = .[[4]][[1]]$ci.lb, heteroRR_upper = .[[4]][[1]]$ci.ub, heteroRR_se = .[[4]][[1]]$se
  ) %>%
  select(., GroupingTerm, heteroCVR:heteroRR_se)

Physiology. <- heterog1 %>%
  filter(., GroupingTerm == "Physiology") %>%
  select(., -data) %>%
  mutate(
    heteroCVR = .[[2]][[1]]$b, heteroCVR_lower = .[[2]][[1]]$ci.lb, heteroCVR_upper = .[[2]][[1]]$ci.ub, heteroCVR_se = .[[2]][[1]]$se,
    heteroVR = .[[3]][[1]]$b, heteroVR_lower = .[[3]][[1]]$ci.lb, heteroVR_upper = .[[3]][[1]]$ci.ub, heteroVR_se = .[[3]][[1]]$se,
    heteroRR = .[[4]][[1]]$b, heteroRR_lower = .[[4]][[1]]$ci.lb, heteroRR_upper = .[[4]][[1]]$ci.ub, heteroRR_se = .[[4]][[1]]$se
  ) %>%
  select(., GroupingTerm, heteroCVR:heteroRR_se)

Metabolism. <- heterog1 %>%
  filter(., GroupingTerm == "Metabolism") %>%
  select(., -data) %>%
  mutate(
    heteroCVR = .[[2]][[1]]$b, heteroCVR_lower = .[[2]][[1]]$ci.lb, heteroCVR_upper = .[[2]][[1]]$ci.ub, heteroCVR_se = .[[2]][[1]]$se,
    heteroVR = .[[3]][[1]]$b, heteroVR_lower = .[[3]][[1]]$ci.lb, heteroVR_upper = .[[3]][[1]]$ci.ub, heteroVR_se = .[[3]][[1]]$se,
    heteroRR = .[[4]][[1]]$b, heteroRR_lower = .[[4]][[1]]$ci.lb, heteroRR_upper = .[[4]][[1]]$ci.ub, heteroRR_se = .[[4]][[1]]$se
  ) %>%
  select(., GroupingTerm, heteroCVR:heteroRR_se)

Morphology. <- heterog1 %>%
  filter(., GroupingTerm == "Morphology") %>%
  select(., -data) %>%
  mutate(
    heteroCVR = .[[2]][[1]]$b, heteroCVR_lower = .[[2]][[1]]$ci.lb, heteroCVR_upper = .[[2]][[1]]$ci.ub, heteroCVR_se = .[[2]][[1]]$se,
    heteroVR = .[[3]][[1]]$b, heteroVR_lower = .[[3]][[1]]$ci.lb, heteroVR_upper = .[[3]][[1]]$ci.ub, heteroVR_se = .[[3]][[1]]$se,
    heteroRR = .[[4]][[1]]$b, heteroRR_lower = .[[4]][[1]]$ci.lb, heteroRR_upper = .[[4]][[1]]$ci.ub, heteroRR_se = .[[4]][[1]]$se
  ) %>%
  select(., GroupingTerm, heteroCVR:heteroRR_se)

Heart. <- heterog1 %>%
  filter(., GroupingTerm == "Heart") %>%
  select(., -data) %>%
  mutate(
    heteroCVR = .[[2]][[1]]$b, heteroCVR_lower = .[[2]][[1]]$ci.lb, heteroCVR_upper = .[[2]][[1]]$ci.ub, heteroCVR_se = .[[2]][[1]]$se,
    heteroVR = .[[3]][[1]]$b, heteroVR_lower = .[[3]][[1]]$ci.lb, heteroVR_upper = .[[3]][[1]]$ci.ub, heteroVR_se = .[[3]][[1]]$se,
    heteroRR = .[[4]][[1]]$b, heteroRR_lower = .[[4]][[1]]$ci.lb, heteroRR_upper = .[[4]][[1]]$ci.ub, heteroRR_se = .[[4]][[1]]$se
  ) %>%
  select(., GroupingTerm, heteroCVR:heteroRR_se)

Eye. <- heterog1 %>%
  filter(., GroupingTerm == "Eye") %>%
  select(., -data) %>%
  mutate(
    heteroCVR = .[[2]][[1]]$b, heteroCVR_lower = .[[2]][[1]]$ci.lb, heteroCVR_upper = .[[2]][[1]]$ci.ub, heteroCVR_se = .[[2]][[1]]$se,
    heteroVR = .[[3]][[1]]$b, heteroVR_lower = .[[3]][[1]]$ci.lb, heteroVR_upper = .[[3]][[1]]$ci.ub, heteroVR_se = .[[3]][[1]]$se,
    heteroRR = .[[4]][[1]]$b, heteroRR_lower = .[[4]][[1]]$ci.lb, heteroRR_upper = .[[4]][[1]]$ci.ub, heteroRR_se = .[[4]][[1]]$se
  ) %>%
  select(., GroupingTerm, heteroCVR:heteroRR_se)

@@

All. <- heterog1_all %>% 
  select(., -data) %>%
  mutate(
    heteroCVR = .[[2]][[1]]$b, heteroCVR_lower = .[[2]][[1]]$ci.lb, heteroCVR_upper = .[[2]][[1]]$ci.ub, heteroCVR_se = .[[2]][[1]]$se,
    heteroVR = .[[3]][[1]]$b, heteroVR_lower = .[[3]][[1]]$ci.lb, heteroVR_upper = .[[3]][[1]]$ci.ub, heteroVR_se = .[[3]][[1]]$se,
    heteroRR = .[[4]][[1]]$b, heteroRR_lower = .[[4]][[1]]$ci.lb, heteroRR_upper = .[[4]][[1]]$ci.ub, heteroRR_se = .[[4]][[1]]$se
  ) %>%
  select(., GroupingTerm, heteroCVR:heteroRR_se)

heterog2 <- bind_rows(Behaviour., Morphology., Metabolism., Physiology., Immunology., Hematology., Heart., Hearing., Eye., All.)
# str(heterog2)

@@

```{r}
sessionInfo() %>% pander()
```

>>>>

Our supplementary material follows the general structure described in our methods and are outlined here in Figure \@ref(fig:FigG1). Generally speaking, the first few sections describe the processes by which the data were cleaned and re-organised for analysis, and how the first order and second order meta-analsyes were conducted. Supplemental tables and figures that are referenecd within the main manuscript can be found at the end of the document or conviently referenecd quickly using the side table of contents panel. We will refer back to this figure as we describe the various aspects of the code and data processing.

@@

### Loading and cleaning data

These functions will load the raw and and do some cleaning to make it ready for further processing and analysis.

```{r}
# loads the raw data, setting some default types for various columns

@@

```{r}
data_subset_parameterid_individual_by_age <- function(mydata, parameter, age_min=0, age_center=100) {
  tmp <- mydata %>%
    filter(age_in_days >= age_min,
           id == parameter) %>%
  # Take results for single individual closest to age_center
    mutate(age_diff = abs(age_center - age_in_days)) %>%
    group_by(biological_sample_id) %>%
    filter(age_diff == min(age_diff)) %>%
    select(-age_diff)

  # Still some individuals with multiple records (because same individual appears under different procedures, so filter to one record)
  j <- match(unique(tmp$biological_sample_id), tmp$biological_sample_id)
  tmp[j, ] 
  }
```

### Population statistics
Create a function called: "calculate_population_stats"
This function groups animals from the same strain and institution together. This is done for each trait separately, and only for traits that have been measured in both sexes. Any group containing fewer than 6 individuals is excluded.

# Function re-organises the data so that the male and female data are side-by-side as columns to make effect size calculations easier.
multi_spread <- function(df, key, value) {
    # quote key
      keyq <- rlang::enquo(key)
    # break value vector into quotes
    valueq <- rlang::enquo(value)
         s <- rlang::quos(!!valueq)

         df                           %>% 
        gather(variable, value, !!!s) %>%
        unite(temp, !!keyq, variable) %>%
        spread(temp, value)
}

# Function will calculate population stats and exclude data that is irrelevant and not useful, it will also just convert the data straight away to wide format making it ready for effect size calculations

### Calculate effect sizes and sample variances
Function: "create_meta_analysis_effect_sizes"

#This function takes the data and computes lnCVR, lnVR and ROM (lnRR) using male and female data to generate effect size statistics for meta-analyses
create_meta_analysis_effect_sizes <- function(mydata, measure = c("CVR", "VR", "ROM")){
  for(i in 1:length(measure)){
    mydata <- metafor::escalc(m1i  = male_x_bar, 
                              m2i  = female_x_bar, 
                              sd1i = male_x_sd, 
                              sd2i = female_x_sd, 
                              n1i  = male_n_ind, 
                              n2i  = female_n_ind, 
                              data = mydata, 
                           measure = measure[i], 
                         var.names = c(paste0("effect_size_", measure[i]), 
                                      paste0("sample_variance_", measure[i])),
                            append = TRUE)
    }
    return(mydata)
}
```

## Load & clean data
We have already done this step and provide a cleaned up file which is less computationally intensive to deal with. The file has been saved in a folder called `export`. However, the `.csv` is provided in case this is preferred and can be imported and processed using following the steps below: [This code is currently excluded from showing in this `.html` file, but can be viewed and enabled / run in the down-loadable .rmd file. However, it requires the .gz file containing the raw data in the "export" folder. Unfortunately, that file can not be provided via Github, as it is too large (274MB). However, it is freely accessible and uploaded to zenodo.org (https://zenodo.org/deposit/3759701)

```{r clean, eval=FALSE}
# Load raw data - save cleaned dataset as RDS for reuse
data_raw <- load_raw(here("data","dr7.0_all_control_data.csv.gz"))
dir.create("export", F, F)

@@

# Meta-analyses
## Population as analysis unit

In this section, we cover the workflow described in Figure \@ref(fig:FigG1) C.

@@

```{r loop, message=FALSE, warning=FALSE}

@@

## Meta-analysis: condensing non-independent traits 
This section describes the workflow depicted in Figure \@ref(fig:FigG1) F.

@@

## Second-order meta-analysis for functional groups
Here, we  describes the workflow depicted in Figure \@ref(fig:FigG1) H.

### Performing meta-analyses 
Three for each of the 9 grouping terms: lnCVR, lnVR, lnRR

#### Preparing the data

@@

# Visualising meta-analysis results
## Preparation for Figure 4 
### Count data, based on First-order meta-analysis results
This includes all separate eligible traits.
Re-ordering of grouping terms. 
Preparation of subplot 4A

@@

### Re-structure data for plotting 
Data are re-structured, and grouping terms are being re-ordered

@@

Panel A shows the numbers of traits across functional groups that are either male-biased (blue-green) or female-biased (orange-red), as calculated in Figure \@ref(fig:FigG1) D. Panel B shows effect sizes and 95% CI from separate meta-analysis for each functional group (step H in Figure \@ref(fig:FigG1)). Both panels represent results evaluated across all traits (Phase 3, Figure \@ref(fig:FigG1)). Traits that are male biased are shown in blue, whereas female bias data is represented in orange.

@@

## Preparation for Figure 5
To further investigate sex bias in this dataset, and in particular if the extent of sex bias differs between traits, we investigate the magnitude of male- and female bias in significantly different traits on (both for means and variability)

@@

#### Figure 5 B - traits with CI not overlapping 0 
# Prepare data 
# create column with 1 = different from zero, 0 = zero included in CI

@@

# as above: Prepare data for traits with CI not overlapping 0
# create column with 1 = different from zero, 0 = zero included in CI

@@

plot3.female.meta.VR <- metacombo_female.plot3.VR %>%
                        mutate(model_lnVR = map(data, ~ metafor::rma.uni(yi = .x$lnVR, sei = (.x$lnVR_upper - .x$lnVR_lower) / (2 * 1.96),
                                                control = list(optimizer = "optim", 
                                                               optmethod = "Nelder-Mead", 
                                                               maxit = 1000), 
                                                verbose = F)))

plot3.female.meta.RR <- metacombo_female.plot3.RR %>%
                        mutate(model_lnRR = map(data, ~ metafor::rma.uni(yi = .x$lnRR, sei = (.x$lnRR_upper - .x$lnRR_lower) / (2 * 1.96),
                                                control = list(optimizer = "optim", 
                                                               optmethod = "Nelder-Mead", 
                                                               maxit = 1000), 
                                                verbose = F)))

@@

plot3.female.meta.CVR.all <- metacombo_female.plot3.CVR.all %>%
                             mutate(model_lnCVR = map(data, ~ metafor::rma.uni(yi = .x$lnCVR, sei=(.x$lnCVR_upper-.x$lnCVR_lower)/(2 * 1.96),
                                                       control = list(optimizer = "optim", 
                                                                      optmethod = "Nelder-Mead", 
                                                                      maxit = 1000),
                                                       verbose = F)))

@@

plot3.female.meta.VR.all <- metacombo_female.plot3.VR.all %>%
  mutate(model_lnVR = map(data, ~ metafor::rma.uni(yi = .x$lnVR, sei = (.x$lnVR_upper - .x$lnVR_lower) / (2 * 1.96),
                          control = list(optimizer = "optim", 
                                         optmethod = "Nelder-Mead", 
                                         maxit = 1000), 
                          verbose = F)))

@@

plot3.female.meta.RR.all <- metacombo_female.plot3.RR.all %>%
                            mutate(model_lnRR = map(data, 
                                                    ~ metafor::rma.uni(yi = .x$lnRR, sei = (.x$lnRR_upper - .x$lnRR_lower) / (2 * 1.96),
                                                    control = list(optimizer = "optim", 
                                                                   optmethod = "Nelder-Mead", 
                                                                   maxit = 1000), 
                                                    verbose = F)))

@@

plot3.female.meta.CVR.b <- as.data.frame(plot3.female.meta.CVR %>% 
                                           group_by(GroupingTerm) %>%
                                          mutate(    lnCVR = map_dbl(model_lnCVR, pluck(2)), 
                                               lnCVR_lower = map_dbl(model_lnCVR, pluck(6)),
                                               lnCVR_upper = map_dbl(model_lnCVR, pluck(7)), 
                                                  lnCVR_se = map_dbl(model_lnCVR, pluck(3)),
                                                  lnCVR_I2 = map_dbl(model_lnCVR, pluck("I2"))))[, c(1, 4:8)]

@@

plot3.female.meta.VR.b <- as.data.frame(plot3.female.meta.VR %>% 
                                        group_by(GroupingTerm) %>%
                                        mutate(      lnVR = map_dbl(model_lnVR, pluck(2)), 
                                               lnVR_lower = map_dbl(model_lnVR, pluck(6)),
                                               lnVR_upper = map_dbl(model_lnVR, pluck(7)), 
                                                  lnVR_se = map_dbl(model_lnVR, pluck(3)),
                                                  lnVR_I2 = map_dbl(model_lnVR, pluck("I2"))))[, c(1, 4:8)]

@@

plot3.female.meta.RR.b <- as.data.frame(plot3.female.meta.RR %>% 
                                        group_by(GroupingTerm) %>%
                                        mutate(      lnRR = map_dbl(model_lnRR, pluck(2)), 
                                               lnRR_lower = map_dbl(model_lnRR, pluck(6)), 
                                               lnRR_upper = map_dbl(model_lnRR, pluck(7)), 
                                                  lnRR_se = map_dbl(model_lnRR, pluck(3)),
                                                  lnRR_I2 = map_dbl(model_lnRR, pluck("I2"))))[, c(1, 4:8)]

@@

lnCVR.ci <- overall3.female.sig %>%
  filter(parameter == "lnCVR") %>%
  mutate(ci.low = lnCVR_lower, 
         ci.high = lnCVR_upper)

# lnVR.ci <- overall3.female.sig  %>% filter(parameter == "lnVR") %>% mutate(ci.low = lnVR_lower, ci.high = lnVR_upper)

lnRR.ci <- overall3.female.sig %>%
  filter(parameter == "lnRR") %>%
  mutate(ci.low = lnRR_lower, ci.high = lnRR_upper)

@@

## Figure 5
Sex biases in trait values across trait groups (Step J in Figure \@ref(fig:FigG1)). The panels in A show differences in the number of affected traits, in variance (lnCVR) and means (lnRR), where there is a significant difference between the sexes (i.e CI not overlapping zero). Panel B shows results for the sex bias in those traits that differ between the sexes (second-order meta-analysis). Triangles represent sex bias in means (response ratio) and black circles differences in the coefficient of variation ratio (mean-adjusted variability). The orange-red bars represent trait groups with a female bias, blue-green bars male-biased traits.

@@

  mutate(
    model_heteroCVR = map(data, ~ metafor::rma.uni(
      yi = .x$var.CVR, sei = sqrt(1 / 2 * (.x$N.CVR - 1)),
      control = list(optimizer = "optim", 
                     optmethod = "Nelder-Mead",
                     maxit = 10000, 
                     stepadj = 0.5), verbose = F)),
    model_heteroVR = map(data, ~ metafor::rma.uni(
      yi = .x$var.VR, sei = sqrt(1 / 2 * (.x$N.VR - 1)),
      control = list(optimizer = "optim", 
                     optmethod = "Nelder-Mead", 
                     maxit = 10000, 
                     stepadj = 0.5), verbose = F)),
    model_heteroRR = map(data, ~ metafor::rma.uni(
      yi = .x$var.RR, sei = sqrt(1 / 2 * (.x$N.RR - 1)),
      control = list(optimizer = "optim", 
                     optmethod = "Nelder-Mead", 
                     maxit = 10000, 
                     stepadj = 0.5), verbose = F)) )

@@

  mutate( model_heteroCVR = map(data, ~ metafor::rma.uni(
      yi = .x$var.CVR, sei = sqrt(1 / 2 * (.x$N.CVR - 1)),
      control = list(optimizer = "optim", 
                     optmethod = "Nelder-Mead", 
                     maxit = 10000, 
                     stepadj = 0.5), verbose = F)),
    model_heteroVR = map(data, ~ metafor::rma.uni(yi = .x$var.VR, sei = sqrt(1 / 2 * (.x$N.VR - 1)),
      control = list(optimizer = "optim", 
                     optmethod = "Nelder-Mead", 
                     maxit = 10000, 
                     stepadj = 0.5), verbose = F)),
    model_heteroRR = map(data, ~ metafor::rma.uni(yi = .x$var.RR, sei = sqrt(1 / 2 * (.x$N.RR - 1)),
      control = list(optimizer = "optim", 
                     optmethod = "Nelder-Mead", 
                     maxit = 10000, 
                     stepadj = 0.5), verbose = F)))

@@

extract_heterogeneity <- function(data, type){

  tmp <- data %>%
         filter(., GroupingTerm == type) %>%
         select(., -data) %>%
         mutate(heteroCVR       = .[[2]][[1]]$b, 
                heteroCVR_lower = .[[2]][[1]]$ci.lb, 
                heteroCVR_upper = .[[2]][[1]]$ci.ub, 
                heteroCVR_se    = .[[2]][[1]]$se,
                heteroVR        = .[[3]][[1]]$b, 
                heteroVR_lower  = .[[3]][[1]]$ci.lb, 
                heteroVR_upper  = .[[3]][[1]]$ci.ub, 
                heteroVR_se     = .[[3]][[1]]$se,
                heteroRR        = .[[4]][[1]]$b, 
                heteroRR_lower  = .[[4]][[1]]$ci.lb, 
                heteroRR_upper  = .[[4]][[1]]$ci.ub, 
                heteroRR_se     = .[[4]][[1]]$se) %>%
        select(., GroupingTerm, heteroCVR:heteroRR_se)

   Behaviour. <- extract_heterogeneity(heterog1, type = "Behaviour")
  Immunology. <- extract_heterogeneity(heterog1, type = "Immunology")
  Hematology. <- extract_heterogeneity(heterog1, type = "Hematology")
     Hearing. <- extract_heterogeneity(heterog1, type = "Hearing")
  Physiology. <- extract_heterogeneity(heterog1, type = "Physiology")
  Metabolism. <- extract_heterogeneity(heterog1, type = "Metabolism")
  Morphology. <- extract_heterogeneity(heterog1, type = "Morphology")
       Heart. <- extract_heterogeneity(heterog1, type = "Heart")
         Eye. <- extract_heterogeneity(heterog1, type = "Eye")

@@

All. <- heterog1_all %>% 
        select(., -data) %>%
        mutate(heteroCVR       = .[[2]][[1]]$b, 
               heteroCVR_lower = .[[2]][[1]]$ci.lb, 
               heteroCVR_upper = .[[2]][[1]]$ci.ub, 
               heteroCVR_se    = .[[2]][[1]]$se, 
               heteroVR        = .[[3]][[1]]$b, 
               heteroVR_lower  = .[[3]][[1]]$ci.lb, 
               heteroVR_upper  = .[[3]][[1]]$ci.ub, 
               heteroVR_se     = .[[3]][[1]]$se, 
               heteroRR        = .[[4]][[1]]$b, 
               heteroRR_lower  = .[[4]][[1]]$ci.lb, 
               heteroRR_upper  = .[[4]][[1]]$ci.ub, 
               heteroRR_se     = .[[4]][[1]]$se) %>%
        select(., GroupingTerm, heteroCVR:heteroRR_se)

heterog2 <- bind_rows(Behaviour., Morphology., Metabolism., Physiology., Immunology., Hematology., Heart., Hearing., Eye., All.)

@@

```{r}
sessionInfo()
```
